---
title: "Import data on R"
---


```{r setup, include=FALSE}

```


# Import data using `read_csv`

1. Load the `readr` package

```{r}

library(readr)
```


2. Read the data "iris.csv" contained in the folder "exercises/data", using the function `read_csv` of  `readr`. Store it in an object (choose the name, just avoid calling it `iris`). Print that object.


```{r}

data_iris <- read_csv("data/iris.csv")
data_iris
iris

```


3. Print `iris` (the classical dataset) directly in the console. 


What are the differences between your dataset and the classical one, when printed  ? 



4. Explore your object using `glimpse` function (load `dplyr` before).


```{r}
library(dplyr)

glimpse(data_iris)
```


### Other arguments of `read_csv`



1. Read the data "cars.csv" contained in the folder "exercises/data", using the function `read_csv` of  `readr`. Store it in an object. Print that object.


```{r}
data_cars <- readr::read_csv("exercises/data/cars.csv")

data_cars
```


Do you see the problem ? Open manually the file and find how to tackle the problem, using other arguments of `read_csv` (see ?read_csv) for details.


```{r}
data_cars <- readr::read_csv("exercises/data/cars.csv", skip = 2)
```

Confirm that you tackled the problem by running `ncol(data_cars)` (replace `data_cars` by the name of the object where you stored the dataset).

There should be 13 columns.


```{r}
ncol(data_cars)
```

Explore your data using `glimpse`
```{r}
glimpse(data_cars)
```

When you see the column "id_car" it is stored as a double, but it may be considered as a character as it is an ID (i.e a name). Use read_csv to consider id_car as character (see argument `col_types` in `read_csv`).


```{r}

# data_cars <- readr::read_csv("exercises/data/cars.csv", skip = 2, col_types = list(col_character(), col_double(), col_double(),col_double(),col_double(),col_double(),col_double(),col_double(),col_double(),col_double(),col_double(),col_double(),   id_car = col_character()))

data_cars <- readr::read_csv("exercises/data/cars.csv", skip = 2, col_types = "cdddddddddddc")


```

In practice, it could be easier to convert the columns into other types after having imported the data.


Ensure you have imported well your data by using `glimpse` again.

```{r}
glimpse(data_cars)
```

# Import data using `readxl`

1. Load the `readxl` package

```{r}
library(readxl)

```


2. Read the data "phenotype.xlsx" contained in the folder "exercises/data", using the function `read_xlsx` of  `readxl`. Store it in an object. Print that object.


```{r}

data_phenotype <- read_xlsx("exercises/data/phenotype.xlsx")

```

3. What happens when you don't specify the sheet argument ? What are the types of the columns ? How can you tackle this problem ?

```{r}
data_phenotype <- read_xlsx("exercises/data/phenotype.xlsx", na = "NOT MEASURED")

```


Read the other sheets of the file and store them in other objects.

```{r}

data_itk <- read_xlsx("exercises/data/phenotype.xlsx", sheet = "itk")
data_index <- read_xlsx("exercises/data/phenotype.xlsx", sheet = "index")


```

The datasets you have created will be useful to us in the next exercises.

