---
title: "Import data on R"
---



# Import data using `read_csv`

1. Load the `readr` package

```{r}


```


2. Read the data "iris.csv" contained in the folder "exercises/data", using the function `read_csv` of  `readr`. Store it in an object (choose the name, just avoid calling it `iris`). Print that object.


```{r}



```


3. Print `iris` (the classical dataset) directly in the console. 


What are the differences between your dataset and the classical one, when printed  ? 




4. Explore your object using `glimpse` function (load `dplyr` before).


```{r}


```


### Other arguments of `read_csv`



1. Read the data "cars.csv" contained in the folder "exercises/data", using the function `read_csv` of  `readr`. Store it in an object. Print that object.


```{r}

```


Do you see the problem (something weird for instance) ? Open manually the file and find how to tackle the problem, using other arguments of `read_csv` (see ?read_csv) for details.


```{r}

```

Confirm that you tackled the problem by running `ncol(data_cars)` (replace `data_cars` by the name of the object where you stored the dataset).

There should be 13 columns.


```{r}

```

Explore your data using `glimpse`
```{r}

```

When you see the column "id_car" it is stored as a double, but it may be considered as a character as it is an ID (i.e a name). Use read_csv to consider id_car as character (see argument `col_types` in `read_csv`, and the `cols`specifications).


```{r}
```

In practice, it could be easier to convert the columns into other types after having imported the data.


Ensure you have imported well your data by using `glimpse` again.

```{r}


```


# Import data using `readxl`

1. Load the `readxl` package

```{r}

```


2. Read the data "phenotype.xlsx" contained in the folder "exercises/data", using the function `read_xlsx` of  `readxl`. Store it in an object. Print that object.


```{r}


```

3. What happens when you don't specify the sheet argument (see `?read_xlsx`)
What are the types of the columns ? Why ?
How can you tackle this problem ?

```{r}

```


Read the other sheets of the file and store them in other objects.

```{r}


```

The datasets you have created will be useful to us in the next exercises.



